---
title: "Data Analysis and Visualization in R"
author: "Jenn Schilling"
date: "June 5, 2022"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      ratio: '16:9'
      img_dir: img
      highlightStyle: github
      highlightLines: true
      slideNumberFormat: "%current%"
      countIncrementalSlides: true
---

```{r setup, include = FALSE}

# Set default chunk options
knitr::opts_chunk$set(echo = TRUE, message = FALSE) 

# Set the default size of figures
knitr::opts_chunk$set(fig.width = 8, fig.height = 5)  

# Load libraries
library(here)
library(explore)
library(tidyverse)
library(scales)

```

```{r xaringan-themer, include = FALSE, warning = FALSE}

style_duo_accent(
  primary_color = "#1c5253",
  secondary_color = "#c8c8c8",
  inverse_background_color = "#1c5253",
  inverse_header_color = "#FFFFFF"
)

```

```{css change-css-options, eval = TRUE, echo = FALSE}

.remark-code{line-height: 1.5; font-size: 60%} # smaller code font size

```


```{r data, include = FALSE}

ipeds_char <- read_csv(here("data", "IPEDS_inst_char.csv"))
ipeds_adm <- read_csv(here("data", "IPEDS_inst_adm.csv"))
ipeds_enrl <- read_csv(here("data", "IPEDS_inst_enrl.csv"))
ipeds_compl <- read_csv(here("data", "IPEDS_inst_compl.csv"))

```

---

class: center, middle, inverse

# Data Exploration

---

class: center, middle, inverse

# Data Analysis

---

class: center, middle, inverse

# Data Visualization

---

# Plotting in R - Tidy Data

```{r plot-data-1, echo = FALSE}

inst_adm_plot <- left_join(inst_list, inst_adm, by = "unitid") %>%
  filter(year == 2019)

sample_n(inst_adm_plot %>% select(instnm, n_applicants, n_admits, n_enrl, sat_score, act_score), size = 10) %>%
  knitr::kable(format = "html")

```

---

# Plotting in R - Data

.pull-left[
```{r plot-1, include = FALSE}

ggplot(data = inst_adm_plot, 
       mapping = aes(x = sat_score,
                     y = act_score)) +
  geom_point()

```

```{r decorate-plot-1a, echo = FALSE}

decorate("plot-1", eval = FALSE) %>%
  flair("data = inst_adm_plot") %>%
  knit_print.with_flair()

```
]

---

# Plotting in R - Aesthetic Mappings

.pull-left[
```{r decorate-plot-1b, echo = FALSE}

decorate("plot-1", eval = FALSE) %>%
  flair("mapping = aes(x = sat_score,
                     y = act_score)") %>%
  knit_print.with_flair()

```
]

---

# Plotting in R - Geometric Objects 

.pull-left[
```{r decorate-plot-1c, echo = FALSE}

decorate("plot-1", eval = FALSE) %>%
  flair("geom_point()") %>%
  knit_print.with_flair()

```
]

---

`r chunk_reveal(chunk_name = "plot-1", title = "#Plotting in R", widths = c(50,50))`

---

# Plotting in R - Scatter Plots

.pull-left[
```{r scatter-2, include = FALSE}

ggplot(data = inst_adm_plot, 
       mapping = aes(x = sat_score,
                     y = act_score,
                     color = instnm)) +
  geom_point()

```

```{r decorate-scatter-2, echo = FALSE}

decorate("scatter-2", eval = FALSE) %>%
  flair("color = instnm") %>%
  knit_print.with_flair()

```

]
.pull-right[
```{r output-scatter-2, ref.label = 'scatter-2', echo = FALSE}

```
]

---

# Plotting in R - Scatter Plots

.pull-left[
```{r scatter-3, include = FALSE}

ggplot(data = inst_adm_plot, 
       mapping = aes(x = sat_score,
                     y = act_score,
                     color = n_enrl)) +
  geom_point()

```

```{r decorate-scatter-3, echo = FALSE}

decorate("scatter-3", eval = FALSE) %>%
  flair("color = n_enrl") %>%
  knit_print.with_flair()

```

]
.pull-right[
```{r output-scatter-3, ref.label = 'scatter-3', echo = FALSE}

```
]

---

# Plotting in R - Scatter Plots
.pull-left[
```{r scatter-4, include = FALSE}

ggplot(data = inst_adm_plot, 
       mapping = aes(x = sat_score,
                     y = act_score,
                     size = n_enrl)) +
  geom_point()

```

```{r decorate-scatter-4, echo = FALSE}

decorate("scatter-4", eval = FALSE) %>%
  flair("size = n_enrl") %>%
  knit_print.with_flair()

```

]
.pull-right[
```{r output-scatter-4, ref.label = 'scatter-4', echo = FALSE}

```
]

---

# Plotting in R - Scatter Plots
.pull-left[
```{r scatter-5, include = FALSE}

ggplot(data = inst_adm_plot, 
       mapping = aes(x = sat_score,
                     y = act_score,
                     size = n_enrl,
                     color = locale)) +
  geom_point()

```

```{r decorate-scatter-5, echo = FALSE}

decorate("scatter-5", eval = FALSE) %>%
  flair("size = n_enrl") %>%
  flair("color = locale") %>%
  knit_print.with_flair()

```

]
.pull-right[
```{r output-scatter-5, ref.label = 'scatter-5', echo = FALSE}

```
]


---

# Plotting in R - New Tidy Data

```{r plot-data-2, echo = FALSE}

inst_adm_plot <- left_join(inst_list, inst_adm, by = "unitid")

inst_adm_plot %>% select(year, instnm, sat_score, n_sat) %>%
  filter(instnm %in% c('Pennsylvania State University-Main Campus',
                       'The University of Texas at Austin',
                       'University of Iowa')) %>%
  arrange(year, instnm) %>%
  knitr::kable(format = "html")

```

---

# Plotting in R - New Geometric Object

.pull-left[
```{r plot-2, include = FALSE}

ggplot(data = inst_adm_plot,
       mapping = aes(x = year,
                     y = sat_score)) +
  geom_line()

```

```{r decorate-plot-2a, echo = FALSE}

decorate("plot-2", eval = FALSE) %>%
  flair("geom_line()") %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-2, ref.label = 'plot-2', echo = FALSE}

```
]

---

# Plotting in R - The `group` Argument

.pull-left[
```{r plot-3, include = FALSE}

ggplot(data = inst_adm_plot,
       mapping = aes(x = year,
                     y = sat_score,
                     group = instnm)) +
  geom_line()

```

```{r decorate-plot-3a, echo = FALSE}

decorate("plot-3", eval = FALSE) %>%
  flair("group = instnm") %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-3, ref.label = 'plot-3', echo = FALSE}

```
]

---

# Plotting in R - Another Line Graph Example

.pull-left[
```{r plot-4, include = FALSE}

ggplot(data = inst_adm_plot,
       mapping = aes(x = year,
                     y = n_sat,
                     group = instnm)) +
  geom_line()

```

```{r decorate-plot-4, echo = FALSE}

decorate("plot-4", eval = FALSE) %>%
  flair("y = n_sat") %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-4, ref.label = 'plot-4', echo = FALSE}

```
]

---

# Plotting in R - Another Line Graph Example

.pull-left[
```{r plot-4a, include = FALSE}

ggplot(data = inst_adm_plot,
       mapping = aes(x = year,
                     y = n_sat,
                     group = instnm,
                     color = instnm)) +
  geom_line()

```

```{r decorate-plot-4a, echo = FALSE}

decorate("plot-4a", eval = FALSE) %>%
  flair("color = instnm") %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-4a, ref.label = 'plot-4a', echo = FALSE}

```
]

---

# Plotting in R - Adding Text

.pull-left[
```{r plot-5, include = FALSE}

ggplot(data = inst_adm_plot,
       mapping = aes(x = year,
                     y = n_sat,
                     group = instnm)) +
  geom_line() +
  geom_text(mapping = aes(label = instnm))

```

```{r decorate-plot-5a, echo = FALSE}

decorate("plot-5", eval = FALSE) %>%
  flair("geom_text(mapping = aes(label = instnm))") %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-5, ref.label = 'plot-5', echo = FALSE}

```
]

---

# Plotting in R - Adding Text for Only 2019

.pull-left[
```{r plot-6, include = FALSE}

ggplot(data = inst_adm_plot,
       mapping = aes(x = year,
                     y = n_sat,
                     group = instnm)) +
  geom_line() +
  geom_text(data = inst_adm_plot %>% 
                   filter(year == 2019),
            mapping = aes(x = year, 
                          y = n_sat, 
                          label = instnm))

```

```{r decorate-plot-6a, echo = FALSE}

decorate("plot-6", eval = FALSE) %>%
  flair("data = inst_adm_plot %>% 
                   filter(year == 2019)") %>%
  flair("mapping = aes(x = year, ") %>%
  flair("y = n_sat, ") %>%
  flair("label = instnm))") %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-6, ref.label = 'plot-6', echo = FALSE}

```
]

---

# Plotting in R - Adding Text for Only 2019

.pull-left[
```{r plot-6b, include = FALSE}

ggplot(data = inst_adm_plot,
       mapping = aes(x = year,
                     y = n_sat,
                     group = instnm)) +
    geom_line() +
    geom_text(data = inst_adm_plot %>% 
                  filter(year == 2019),
              mapping = aes(x = year,
                            y = n_sat,
                            label = instnm),
              hjust = 0)

```

```{r decorate-plot-6b, echo = FALSE}

decorate("plot-6b", eval = FALSE) %>%
  flair("hjust = 0)") %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-6b, ref.label = 'plot-6b', echo = FALSE}

```
]

---

# Plotting in R - Adding Text for Only 2019

.pull-left[
```{r plot-6c, include = FALSE}

ggplot(data = inst_adm_plot,
       mapping = aes(x = year,
                     y = n_sat,
                     group = instnm)) +
    geom_line() +
    geom_text(data = inst_adm_plot %>% 
                  filter(year == 2019),
              mapping = aes(x = year,
                            y = n_sat,
                            label = instnm),
              hjust = 0) + 
    scale_x_continuous(limits = c(2017, 2020.5))

```

```{r decorate-plot-6c, echo = FALSE}

decorate("plot-6c", eval = FALSE) %>%
  flair("scale_x_continuous(limits = c(2017, 2020.5))") %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-6c, ref.label = 'plot-6c', echo = FALSE}

```
]

---

```{r plot-7, include = FALSE}

ggplot(data = inst_adm_plot,
       mapping = aes(x = year,
                     y = n_sat,
                     group = instnm)) +
  geom_line() +
  geom_text(data = inst_adm_plot %>% 
                   filter(year == 2019),
            mapping = aes(x = Inf,
                          y = n_sat,
                          label = instnm),
            hjust = -.01) +
  coord_cartesian(clip = "off") +
  theme(plot.margin = margin(0, 250, 0, 0))

```

`r chunk_reveal(chunk_name = "plot-7", title = "#Plotting in R - Format Text Labels", widths = c(50,50))`

---

class: center, middle, inverse

# Day 2

---

# Plotting in R - Tidy Data

```{r plot-data-3, echo = FALSE}

sample_n(inst_list %>% select(unitid, instnm, locale), size = 10) %>%
  knitr::kable(format = "html")

```

---

# Plotting in R - Bar Graph

.pull-left[
```{r plot-8, include = FALSE}

ggplot(data = inst_list,
       mapping = aes(x = locale)) +
  geom_bar()

```

```{r decorate-plot-8a, echo = FALSE}

decorate("plot-8", eval = FALSE) %>%
  flair("x = locale") %>%
  flair("geom_bar()") %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-8, ref.label = 'plot-8', echo = FALSE}

```
]

---

# Plotting in R - Bar Graph Sorted

.pull-left[
```{r plot-9, include = FALSE}

ggplot(data = inst_list,
       mapping = aes(x = fct_infreq(locale))) +
  geom_bar()

```

```{r decorate-plot-9a, echo = FALSE}

decorate("plot-9", eval = FALSE) %>%
  flair("fct_infreq(locale))") %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-9, ref.label = 'plot-9', echo = FALSE}

```
]

---

# Plotting in R - Tidy Data

```{r plot-data-4, echo = FALSE}

inst_enrlmnt_plot <- left_join(inst_list, inst_enrlmnt, by = "unitid")

sample_n(inst_enrlmnt_plot %>% select(year, instnm, locale, lstudy, gender, n_enrl), size = 10) %>%
  knitr::kable(format = "html")

```

---

# Plotting in R - Bar Graph Not Counts

.pull-left[
```{r plot-10, include = FALSE}

ggplot(data = inst_enrlmnt_plot %>% 
         filter(lstudy == "All students", 
                gender == "Total",
                year == 2019),
       mapping = aes(x = locale,
                     y = n_enrl)) +
  geom_col()

```

```{r decorate-plot-10a, echo = FALSE}

decorate("plot-10", eval = FALSE) %>%
  flair("x = locale") %>%
  flair("y = n_enrl") %>%
  flair("geom_col()") %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-10, ref.label = 'plot-10', echo = FALSE}

```
]

---

# Plotting in R - What is this actually doing?

.pull-left[
```{r plot-11, include = FALSE}

ggplot(data = inst_enrlmnt_plot %>% 
         filter(lstudy == "All students", 
                gender == "Total",
                year == 2019),
       mapping = aes(x = locale,
                     y = n_enrl,
                     fill = instnm)) +
  geom_col()

```

```{r decorate-plot-11a, echo = FALSE}

decorate("plot-11", eval = FALSE) %>%
  flair("fill = instnm") %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-11, ref.label = 'plot-11', echo = FALSE}

```
]

---

# Plotting in R - Tidy (Aggregated) Data

.pull-left[
```{r plot-data-5}

inst_enrlmnt_plot_sub <- inst_enrlmnt_plot %>%
  filter(lstudy == "All students", 
         gender == "Total",
         year == 2019) %>%
  group_by(locale) %>%
  summarize(med_enrl = median(n_enrl),
            .groups = "drop")

```
]

.pull-right[
```{r plot-data-5-show, echo = FALSE}
inst_enrlmnt_plot_sub %>% 
           select(locale, med_enrl) %>%
  knitr::kable(format = "html")
```

]

---

# Plotting in R - Bar Graph Aggregate

.pull-left[
```{r plot-12, include = FALSE}

ggplot(data = inst_enrlmnt_plot_sub,
       mapping = aes(x = locale,
                     y = med_enrl)) +
  geom_col()

```

```{r decorate-plot-12a, echo = FALSE}

decorate("plot-12", eval = FALSE) %>%
  flair("data = inst_enrlmnt_plot_sub") %>%
  flair("y = med_enrl") %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-12, ref.label = 'plot-12', echo = FALSE}

```
]

---

# Plotting in R - Bar Graph Aggregate Sorted

.pull-left[
```{r plot-13, include = FALSE}

ggplot(data = inst_enrlmnt_plot_sub,
       mapping = aes(x = reorder(locale,
                                 -med_enrl),
                     y = med_enrl)) +
  geom_col()

```

```{r decorate-plot-13a, echo = FALSE}

decorate("plot-13", eval = FALSE) %>%
  flair("reorder(locale,") %>%
  flair("-med_enrl),") %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-13, ref.label = 'plot-13', echo = FALSE}

```
]

---

```{r plot-14, include = FALSE}

ggplot(data = inst_enrlmnt_plot %>% 
         filter(lstudy == "All students", 
                gender == "Total",
                year == 2019),
       mapping = aes(x = n_enrl,
                     y = reorder(instnm, 
                                 n_enrl))) +
  geom_col() +
  facet_wrap(~ locale)

```


`r chunk_reveal(chunk_name = "plot-14", title = "# Plotting in R - Facet Wrap", widths = c(50,50))`

---

# Plotting in R - Facet Wrap Formatting

.pull-left[
```{r plot-15, include = FALSE}

ggplot(data = inst_enrlmnt_plot %>% 
         filter(lstudy == "All students", 
                gender == "Total",
                year == 2019),
       mapping = aes(x = n_enrl,
                     y = reorder(instnm, 
                                 n_enrl))) +
  geom_col() +
  facet_wrap(~ locale,
             scales = "free_y",
             ncol = 1)

```

```{r decorate-plot-15a, echo = FALSE}

decorate("plot-15", eval = FALSE) %>%
  flair('scales = "free_y"') %>%
  flair("ncol = 1") %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-15, ref.label = 'plot-15', echo = FALSE}

```
]


---

# Plotting in R - Tidy Data

```{r plot-data-6, echo = FALSE}

inst_enrlmnt_plot_sub <- inst_enrlmnt_plot %>%
  filter(lstudy == "All students", gender == "Total") %>%
  group_by(year, locale) %>%
  summarize(med_enrl = median(n_enrl),
            .groups = "drop")

sample_n(inst_enrlmnt_plot_sub %>% select(year, locale, med_enrl), size = 10) %>%
  knitr::kable(format = "html")

```

---

# Plotting in R - Adding Color & Position Dodge

.pull-left[
```{r plot-16, include = FALSE}

ggplot(data = inst_enrlmnt_plot_sub,
       mapping = aes(x = year,
                     y = med_enrl,
                     fill = locale)) +
  geom_col(position = "dodge") 

```

```{r decorate-plot-16a, echo = FALSE}

decorate("plot-16", eval = FALSE) %>%
  flair('fill = locale') %>%
  flair('position = "dodge"') %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-16, ref.label = 'plot-16', echo = FALSE}

```
]

---

# Plotting in R - Facet & Remove Legend

.pull-left[
```{r plot-17, include = FALSE}

ggplot(data = inst_enrlmnt_plot_sub,
       mapping = aes(x = year,
                     y = med_enrl,
                     fill = locale)) +
  geom_col(position = "dodge") +
  facet_wrap(~ locale) +
  guides(fill = FALSE)

```

```{r decorate-plot-17a, echo = FALSE}

decorate("plot-17", eval = FALSE) %>%
  flair('facet_wrap(~ locale)') %>%
  flair('guides(fill = FALSE)') %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-17, ref.label = 'plot-17', echo = FALSE}

```
]

---

# Plotting in R - Show with Lines Instead

.pull-left[
```{r plot-18, include = FALSE}

ggplot(data = inst_enrlmnt_plot_sub,
       mapping = aes(x = year,
                     y = med_enrl,
                     fill = locale)) +
  geom_line() +
  facet_wrap(~ locale) +
  guides(fill = FALSE) 

```

```{r decorate-plot-18a, echo = FALSE}

decorate("plot-18", eval = FALSE) %>%
  flair('geom_line()') %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-18, ref.label = 'plot-18', echo = FALSE}

```
]

---

# Plotting in R - Map Color

.pull-left[
```{r plot-19, include = FALSE}

ggplot(data = inst_enrlmnt_plot_sub,
       mapping = aes(x = year,
                     y = med_enrl,
                     color = locale)) +
  geom_line() +
  facet_wrap(~ locale) +
  guides(color = FALSE) 

```

```{r decorate-plot-19a, echo = FALSE}

decorate("plot-19", eval = FALSE) %>%
  flair('color') %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-19, ref.label = 'plot-19', echo = FALSE}

```
]

---

# Plotting in R - Tidy Data

```{r plot-data-7, echo = FALSE}

inst_compl_plot <- left_join(inst_list, inst_compl, by = "unitid") 

sample_n(inst_compl_plot %>% select(year, unitid, instnm, awlevelc, gender, n_deg), size = 10) %>%
  arrange(year, unitid, awlevelc, gender) %>%
  knitr::kable(format = "html")

```

---

# Plotting in R - Starting Plot

.pull-left[
```{r plot-20, eval = FALSE}

ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", 
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = awlevelc)) +
  geom_line()

```
]

.pull-right[
```{r output-plot-20, ref.label = 'plot-20', echo = FALSE}

```
]

---

# Plotting in R - Reorder Degree Levels

.pull-left[
```{r plot-21, include = FALSE}

ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", 
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line()

```

```{r decorate-plot-21a, echo = FALSE}

decorate("plot-21", eval = FALSE) %>%
  flair('factor(awlevelc,
                      levels = c("Bachelor\'s degree",
                                 "Master\'s degree",
                                 "Doctor\'s degree"))))') %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-21, ref.label = 'plot-21', echo = FALSE}

```
]


---

# Plotting in R - Add Titles

.pull-left[
```{r plot-22, include = FALSE}

ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", 
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line() +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "")

```

```{r decorate-plot-22a, echo = FALSE}

decorate("plot-22", eval = FALSE) %>%
  flair('labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "")') %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-22, ref.label = 'plot-22', echo = FALSE}

```
]

---

# Plotting in R - Move Legend

.pull-left[
```{r plot-23, include = FALSE}

ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", 
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line() +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "") +
  theme(legend.position = "top")

```

```{r decorate-plot-23a, echo = FALSE}

decorate("plot-23", eval = FALSE) %>%
  flair('theme(legend.position = "top")') %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-23, ref.label = 'plot-23', echo = FALSE}

```
]

---

# Plotting in R - Add Caption

.pull-left[
```{r plot-24, include = FALSE}

ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", 
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line() +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme(legend.position = "top")

```

```{r decorate-plot-24a, echo = FALSE}

decorate("plot-24", eval = FALSE) %>%
  flair('caption = "Data from IPEDS Completions Survey"') %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-24, ref.label = 'plot-24', echo = FALSE}

```
]

---

# Plotting in R - Add Theme

.pull-left[
```{r plot-25, include = FALSE}

ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", 
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line() +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme_minimal() +
  theme(legend.position = "top")

```

```{r decorate-plot-25a, echo = FALSE}

decorate("plot-25", eval = FALSE) %>%
  flair('theme_minimal()') %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-25, ref.label = 'plot-25', echo = FALSE}

```
]

---

# Plotting in R - Add Commas to Y-Axis

.pull-left[
```{r plot-26, include = FALSE}

ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", 
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme_minimal() +
  theme(legend.position = "top")

```

```{r decorate-plot-26a, echo = FALSE}

decorate("plot-26", eval = FALSE) %>%
  flair('scale_y_continuous(labels = scales::comma)') %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-26, ref.label = 'plot-26', echo = FALSE}

```
]

---

# Plotting in R - Remove Decimals from X-Axis

.pull-left[
```{r plot-27, include = FALSE}

ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", 
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = c(2017, 2018, 2019)) +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme_minimal() +
  theme(legend.position = "top")

```

```{r decorate-plot-27a, echo = FALSE}

decorate("plot-27", eval = FALSE) %>%
  flair('scale_x_continuous(breaks = c(2017, 2018, 2019))') %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-27, ref.label = 'plot-27', echo = FALSE}

```
]

---

# Plotting in R - Add Academic Year Labels

.pull-left[
```{r plot-28, include = FALSE}

ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", 
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = c(2017, 2018, 2019),
                     labels = c("2016-17", 
                                "2017-18", 
                                "2018-19")) +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme_minimal() +
  theme(legend.position = "top")

```

```{r decorate-plot-28a, echo = FALSE}

decorate("plot-28", eval = FALSE) %>%
  flair('labels = c("2016-17",') %>%
  flair('"2017-18",') %>%
  flair('"2018-19")') %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-28, ref.label = 'plot-28', echo = FALSE}

```
]

---

# Plotting in R - Add Points on Top of Lines

.pull-left[
```{r plot-29, include = FALSE}

ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total",
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line(show.legend = FALSE) +
  geom_point() +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = c(2017, 2018, 2019),
                     labels = c("2016-17", 
                                "2017-18", 
                                "2018-19")) +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme_minimal() +
  theme(legend.position = "top")

```

```{r decorate-plot-29a, echo = FALSE}

decorate("plot-29", eval = FALSE) %>%
  flair('show.legend = FALSE') %>%
  flair('geom_point()') %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-29, ref.label = 'plot-29', echo = FALSE}

```
]

---

# Plotting in R - Change Color with ColorBrewer

.pull-left[
```{r plot-30, include = FALSE}

ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total",
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line(show.legend = FALSE) +
  geom_point() +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = c(2017, 2018, 2019),
                     labels = c("2016-17", 
                                "2017-18", 
                                "2018-19")) +
  scale_color_brewer(palette = "Dark2") +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme_minimal() +
  theme(legend.position = "top")

```

```{r decorate-plot-30a, echo = FALSE}

decorate("plot-30", eval = FALSE) %>%
  flair('scale_color_brewer(palette = "Dark2")') %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-30, ref.label = 'plot-30', echo = FALSE}

```
]

---

# Plotting in R - Change Color Manually

.pull-left[
```{r plot-31, include = FALSE}

ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total",
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line(show.legend = FALSE) +
  geom_point() +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = c(2017, 2018, 2019),
                     labels = c("2016-17", 
                                "2017-18", 
                                "2018-19")) +
  scale_color_manual(values = c("#0C234B", 
                                "#AB0520", 
                                "#378DBD"))  +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme_minimal() +
  theme(legend.position = "top")

```

```{r decorate-plot-31a, echo = FALSE}

decorate("plot-31", eval = FALSE) %>%
  flair('scale_color_manual(values = c("#0C234B",') %>% 
  flair('"#AB0520",') %>%
  flair('"#378DBD"))') %>%
  knit_print.with_flair()

```
]

.pull-right[
```{r output-plot-31, ref.label = 'plot-31', echo = FALSE}

```
]

---

# Plotting in R -Saving

.pull-left[
```{r plot-32, include = FALSE}

line_plot <- ggplot(
  data = inst_compl_plot %>% 
         filter(gender == "Total",
                unitid == 104179),
  mapping = aes(x = year,
                y = n_deg,
                color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line(show.legend = FALSE) +
  geom_point() +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = c(2017, 2018, 2019),
                     labels = c("2016-17", "2017-18","2018-19")) +
  scale_color_manual(values = c("#0C234B", "#AB0520", "#378DBD"))  +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme_minimal() +
  theme(legend.position = "top")

```

```{r decorate-plot-32a, echo = FALSE}

decorate("plot-32", eval = FALSE) %>%
  flair('line_plot <-') %>% 
  knit_print.with_flair()

```
]

.pull-right[
```{r save-plot-32, eval = FALSE, include = FALSE}

ggsave(filename = here("plots", "line_plot.png"),
       plot = line_plot,
       device = "png",
       width = 8,
       height = 5,
       type = "cairo")

```

```{r decorate-save-plot-32a, echo = FALSE}

decorate("save-plot-32", eval = FALSE) %>%
  flair('ggsave') %>% 
  knit_print.with_flair()

```
]