---
title: "Data Analysis and Visualization in R"
author: "Jenn Schilling"
date: "June 5, 2022"
output:
 html_document:
  theme: "flatly" 
  toc: TRUE
  toc_float: TRUE
---


This file contains contains code for pre-conference AIR Forum Workshop: Data Analysis Visualization in R, presented for the Association for Institutional Research by Jenn Schilling on June 5, 2022. 

**Materials developed by Jenn Schilling.**  

# Setup 

```{r setup, message = FALSE, warning = FALSE}

# Set default chunk options
knitr::opts_chunk$set(echo = TRUE, message = FALSE) 

# Set the default size of figures
knitr::opts_chunk$set(fig.width = 8, fig.height = 5)  

# Load libraries
library(here)
library(explore)
library(tidyverse)
library(scales)

```


# Read the Data

```{r read-data}

# Data from IPEDS
ipeds_char <- read_csv(here("data", "IPEDS_inst_char.csv"))
ipeds_adm <- read_csv(here("data", "IPEDS_inst_adm.csv"))
ipeds_enrl <- read_csv(here("data", "IPEDS_inst_enrl.csv"))
ipeds_compl <- read_csv(here("data", "IPEDS_inst_compl.csv"))

```



# Data Exploration

```{r explore-data, eval = FALSE}

# Interactive exploration of the data
explore(ipeds_adm)

# Create a report for exploratory data analysis
report(ipeds_adm,
       output_file = "ipeds_adm_report.html",
       output_dir = here("output"))

# Create a report for exploratory data analysis with a target variable
report(ipeds_adm,
       output_file = "ipeds_adm_report_target.html",
       output_dir = here("output"),
       target = n_enrl)

```


## Data Exploration - Your Turn

Use the `explore()` or `report()` to explore one of the other IPEDS datasets.

What do you notice? 

Are there any outliers or unusual patterns in any variable?

Are there any missing values? 

Does the shape of the data reveal anything interesting? 

Learn more at the package website: https://github.com/rolkra/explore


```{r explore-data-your-turn}




```



# Data Analysis

```{r analyze-data}

## Summarize Data ##

# General, base summary of every column/variable
summary(ipeds_adm)

# Compute summary statistics for each institution 
ipeds_adm_summary <- ipeds_adm %>%
  group_by(instnm) %>%
  summarize(median_applicants = median(n_applicants, na.rm = TRUE),
            mean_applicants = mean(n_applicants, na.rm = TRUE),
            std_applicants = sd(n_applicants, na.rm = TRUE),
            max_applicants = max(n_applicants, na.rm = TRUE),
            min_applicants = min(n_applicants, na.rm = TRUE),
            num_years = n()) %>%
  ungroup()
  
View(ipeds_adm_summary)

## Manipulate Data ##

# Add a new column
ipeds_adm_new_col <- ipeds_adm %>%
  mutate(yield = n_admits / n_applicants)

View(ipeds_adm_new_col)

# Filter data to get rows of interest
ipeds_adm_subset_rows <- ipeds_adm %>%
  filter(year >= 2019)

View(ipeds_adm_subset_rows)

# Select only columns of interest
ipeds_adm_subset_col <- ipeds_adm %>%
  select(year, unitid, instnm, n_applicants, n_sat, sat_score, n_act, act_score)

View(ipeds_adm_subset_col)

# Add more columns/variables by joining
View(ipeds_adm)
View(ipeds_char)

ipeds_adm_char <- left_join(ipeds_adm, ipeds_char, by = c("unitid", "instnm"))

View(ipeds_adm_char)


## Correlation ##
ipeds_adm_corr <- cor(ipeds_adm$n_admits, ipeds_adm$n_enrl)

ipeds_adm_corr

## Linear Regression ##
ipeds_adm_model <- lm(n_enrl ~ n_admits, data = ipeds_adm)

summary(ipeds_adm_model)


```


## Data Analysis - Your Turn

Use the `summarize()` function to create summary statistics for a different column or different IPEDS dataset. 

Try to add a new column with the `mutate()` function or subset the data using `filter()` or `select()` in one of the IPEDS datasets.

Put two IPEDS datasets together using a `left_join()`, `right_join()`, or `inner_join()`.

Explore a linear regression using `corr()` and `lm()`.

Learn more at the package website: https://dplyr.tidyverse.org/

If your data is not already in tidy format, the {tidyr} package provides useful functions for processing data. Learn more at the package website: https://tidyr.tidyverse.org/


```{r analyze-data-your-turn}




```



# Data Visualization

## Scatterplot

```{r data-viz-01-scatter}

# Is there a relationship between SAT score and ACT score in Fall 2019?

# Get the data ready for the plot
ipeds_adm_plot <- ipeds_adm %>%
  filter(year == 2019)

# Step 1: Specify the data
ggplot(data = ipeds_adm_plot)

# Step 2: Specify the aesthetic mapping
ggplot(data = ipeds_adm_plot,
       mapping = aes(x = sat_score,
                     y = act_score))

# Step 3: Specify the geometric object
ggplot(data = ipeds_adm_plot,
       mapping = aes(x = sat_score,
                     y = act_score)) +
  geom_point()

# Step 4: Add to aesthetic mapping

# Categorical color
ggplot(data = ipeds_adm_plot,
       mapping = aes(x = sat_score,
                     y = act_score,
                     color = instnm)) +
  geom_point()

# Sequential color
ggplot(data = ipeds_adm_plot,
       mapping = aes(x = sat_score,
                     y = act_score,
                     color = n_enrl)) +
  geom_point()

# Size
ggplot(data = ipeds_adm_plot,
       mapping = aes(x = sat_score,
                     y = act_score,
                     size = n_enrl)) +
  geom_point()


```

## Line Graph

```{r data-viz-02-line}

# How does the number of students who submit an SAT score change over time?

# This time, we will use the whole ipeds_adm dataset

# Set up the data, aesthetic mapping, and geometric objects
ggplot(data = ipeds_adm,
       mapping = aes(x = year,
                     y = n_sat)) +
  geom_line()

# Something doesn't look right - there is only one line

# Need to add the group argument in the aes() mapping to tell R to create one line per group (institution)
ggplot(data = ipeds_adm,
       mapping = aes(x = year,
                     y = n_sat,
                     group = instnm)) +
  geom_line()

# How can we identify the outlier?

# Option 1: Add Color
ggplot(data = ipeds_adm,
       mapping = aes(x = year,
                     y = n_sat,
                     group = instnm,
                     color = instnm)) +
  geom_line()

# Option 2: Add Labels
# 2a: Add label for every data point
ggplot(data = ipeds_adm,
       mapping = aes(x = year,
                     y = n_sat,
                     group = instnm,
                     label = instnm)) +
  geom_line() +
  geom_text() 

# 2b: Add label for only the last data point

# We can set data and mapping arguments for individual geometric objects

# First, let's make a new data table with only the labels 
ipeds_adm_labels <- ipeds_adm %>%
  filter(!is.na(n_sat)) %>% # remove NA values for n_sat
  group_by(instnm) %>% # group the data by institution
  summarise(year = max(year)) %>% # find the maximum year
  ungroup() %>% # remove grouping
  left_join(ipeds_adm, by = c("instnm", "year")) %>% # join to original dataset to get the n_sat value for that year
  select(instnm, year, n_sat) # keep only the columns we need
  
# Plot
ggplot(data = ipeds_adm,
       mapping = aes(x = year,
                     y = n_sat,
                     group = instnm)) +
  geom_line() +
  geom_text(data = ipeds_adm_labels,
            mapping = aes(x = year,
                          y = n_sat,
                          label = instnm))

# Format labels and allow them to show outside the plot area
ggplot(data = ipeds_adm,
       mapping = aes(x = year,
                     y = n_sat,
                     group = instnm)) +
  geom_line() +
  geom_text(data = ipeds_adm_labels,
            mapping = aes(x = year, 
                          y = n_sat,
                          label = instnm),
            hjust = 0) +
  coord_cartesian(clip = "off") + # show the labels even though they are outside the plot area
  theme(plot.margin = margin(0, 250, 0, 0)) # expand the plot margin so we can see the labels

# Format labels so they do not overlap
library(ggrepel)
ggplot(data = ipeds_adm,
       mapping = aes(x = year,
                     y = n_sat,
                     group = instnm)) +
  geom_line() +
  geom_text_repel(data = ipeds_adm_labels,
                  mapping = aes(x = year, 
                                y = n_sat,
                                label = instnm)) 

```

## Bar Graph

```{r data-viz-03-bar}

# How many institutions are in each type of locale?

# Set up the data, aesthetic mapping, and geometric objects
ggplot(data = ipeds_char,
       mapping = aes(x = locale)) +
  geom_bar()

# geom_bar() computes counts for us - we just need to provide the x-axis category

# Can we sort them?
# Yes, by using the fct_infreq() function from the forcats package in tidyverse
# fct_infreq() reorders the levels by their frequencies
ggplot(data = ipeds_char,
       mapping = aes(x = fct_infreq(locale))) +
  geom_bar()

# What if we want to make a bar plot with something other than counts?

# Create a new dataset with enrollment data and institutional characteristics
inst_enrlmnt_data <- left_join(ipeds_char, ipeds_enrl, by = c("unitid", "instnm"))

View(inst_enrlmnt_data)

# What does enrollment look like by locale?

# Plot total enrollment of all students in Fall 2019 by locale
ggplot(data = inst_enrlmnt_data %>% 
         filter(lstudy == "All students", gender == "Total", year == 2019),
       mapping = aes(x = locale,
                     y = n_enrl)) +
  geom_col() # geom_col() plots a bar with the data specified in the y argument of aes()

# Let's see what's actually happening in this plot - the enrollments for each institution are stacked on top of one another
ggplot(data = inst_enrlmnt_data %>% 
         filter(lstudy == "All students", gender == "Total", year == 2019),
       mapping = aes(x = locale,
                     y = n_enrl,
                     fill = instnm)) + # fill adds color to a bar plot
  geom_col()

# What if we want the median enrollment in each area?
# We can use filter(), group_by(), and summarize() to compute an aggregation
inst_enrlmnt_data_sub <- inst_enrlmnt_data %>%
  filter(lstudy == "All students", gender == "Total", year == 2019) %>%
  group_by(locale) %>%
  summarize(med_enrl = median(n_enrl)) %>%
  ungroup()

View(inst_enrlmnt_data_sub)

ggplot(data = inst_enrlmnt_data_sub,
       mapping = aes(x = locale,
                     y = med_enrl)) +
  geom_col()

# Now plot the aggregated data and arrange the bars in order of median enrollemtn
# This time, use the reorder() function to arrange the locale column by the med_enrl column because we have a numeric value we want to use for the ordering
ggplot(data = inst_enrlmnt_data_sub,
       mapping = aes(x = reorder(locale, -med_enrl),
                     y = med_enrl)) +
  geom_col()


```

## Pie Chart

```{r data-viz-04-pie}

# What proportion of undergraduate enrollment at the University of Arizona in Fall 2020 was men vs. women?

# Get the subset of the data we're interested in
ua_enrl <- ipeds_enrl %>%
  filter(unitid == 104179 & year == 2020 & gender != "Total" & lstudy == "Undergraduate")

View(ua_enrl)

# Plot the data, aesthetics mapping, and geometric object
ggplot(data = ua_enrl,
       mapping = aes(x = year,
                     y = n_enrl,
                     fill = gender)) +
  geom_col()

# Change the coordinate system to make a pie chart
ggplot(data = ua_enrl,
       mapping = aes(x = year,
                     y = n_enrl,
                     fill = gender)) +
  geom_col() +
  coord_polar(theta = "y")

# What if we want to use percentages?

# Make a new data table with the percentages
ua_enrl_perc <- ua_enrl %>%
  mutate(p_enrl = n_enrl / sum(n_enrl))

View(ua_enrl_perc)

# Plot the percentages
ggplot(data = ua_enrl_perc,
       mapping = aes(x = year,
                     y = p_enrl,
                     fill = gender)) +
  geom_col() +
  coord_polar(theta = "y",
              start = 0) 

```

## Faceting


```{r data-viz-05-facets}

# What does enrollment look like by institution and locale?

# Plot the data, aesthetic mapping, and geometric object
ggplot(data = inst_enrlmnt_data %>% 
         filter(lstudy == "All students", 
                gender == "Total",
                year == 2019),
       mapping = aes(x = n_enrl,
                     y = reorder(instnm, n_enrl))) +
  geom_col()

# We could add color to see the locale
ggplot(data = inst_enrlmnt_data %>% 
         filter(lstudy == "All students", 
                gender == "Total",
                year == 2019),
       mapping = aes(x = n_enrl,
                     y = reorder(instnm, n_enrl),
                     fill = locale)) +
  geom_col()


# Or we could make small charts for each locale
ggplot(data = inst_enrlmnt_data %>% 
         filter(lstudy == "All students", 
                gender == "Total",
                year == 2019),
       mapping = aes(x = n_enrl,
                     y = reorder(instnm, n_enrl))) +
  geom_col() +
  facet_wrap(~ locale) # Make a small plot for each locale

# Make some formatting changes for the facets
ggplot(data = inst_enrlmnt_data %>% 
         filter(lstudy == "All students", 
                gender == "Total",
                year == 2019),
       mapping = aes(x = n_enrl,
                     y = reorder(instnm, n_enrl))) +
  geom_col() +
  facet_wrap(~ locale,
             ncol = 1, # Put all the plots in one column
             scales = "free_y") # Let the y-axis be "free" so only the relevant institutions to that locale show


```

## Facets with Bars and Lines

```{r data-viz-06-bar}

# What does median enrollment look like over time in each locale?

# Make a sub dataset for the plot
inst_enrlmnt_data_sub <- inst_enrlmnt_data %>%
  filter(lstudy == "All students", gender == "Total") %>%
  group_by(year, locale) %>%
  summarize(med_enrl = median(n_enrl),
            .groups = "drop")

View(inst_enrlmnt_data_sub)

# Plot - this creates a stacked bar graph
ggplot(data = inst_enrlmnt_data_sub,
       mapping = aes(x = year,
                     y = med_enrl,
                     fill = locale)) +
  geom_col()

# Plot - this creates a side-by-side bar graph
ggplot(data = inst_enrlmnt_data_sub,
       mapping = aes(x = year,
                     y = med_enrl,
                     fill = locale)) +
  geom_col(position = "dodge")

# Make a small plot for each locale
ggplot(data = inst_enrlmnt_data_sub,
       mapping = aes(x = year,
                     y = med_enrl,
                     fill = locale)) +
  geom_col(position = "dodge") +
  facet_wrap(~ locale)

# Remove the legend since we no longer need it
ggplot(data = inst_enrlmnt_data_sub,
       mapping = aes(x = year,
                     y = med_enrl,
                     fill = locale)) +
  geom_col(position = "dodge") +
  facet_wrap(~ locale) +
  guides(fill = "none")

# Since we're showing data over time, maybe a line graph would be better
# Can we just replace geom_col() with geom_line()?
ggplot(data = inst_enrlmnt_data_sub,
       mapping = aes(x = year,
                     y = med_enrl,
                     fill = locale)) +
  geom_line() +
  facet_wrap(~ locale) +
  guides(fill = "none")

# Where did the color go? Also, notice the change in the y-axis - it no longer starts at 0

# Replace "fill" with "color" to get the line color back
ggplot(data = inst_enrlmnt_data_sub,
       mapping = aes(x = year,
                     y = med_enrl,
                     color = locale)) +
  geom_line() +
  facet_wrap(~ locale) +
  guides(color = "none")

# Bars, histograms, and boxplots use fill; lines and points use color

```


## Formatting and Saving

```{r data-viz-07-format}

# So far, we’ve gone over a few different plot types, how to add color, and faceting. But how do you edit the colors and theme of the plot to get something that looks really nice? We’ve seen some pretty strange and unhelpful axis labels? How do we fix that?

# What do degree completions look like over time at the University of Arizona?

# Create dataset for the plot
ua_compl <- ipeds_compl %>%
  filter(gender == "Total", unitid == 104179)

# Plot - data, aesthetic mapping, geometric object
# Here we are also assigning the plot to an object p to make it easier to add on to
p <- ggplot(data = ua_compl,
       mapping = aes(x = year,
                     y = n_deg,
                     color = awlevelc)) +
  geom_line()

p

# Things we want to fix:
# 1. Degree levels are not in a logical order
# 2. X-axis labels need to include the whole academic year
# 3. Axis titles are not very helpful
# 4. There's not plot title or data source noted
# 5. Use colors other than the R defaults for the lines
# 6. Remove the grey background from the plot

# 1. Reorder the degree levels
# The easiest way to do this is in the data for the plot
ua_compl <- ipeds_compl %>%
  filter(gender == "Total", unitid == 104179) %>%
  mutate(awlevelc = factor(awlevelc,
                           levels = c("Bachelor's degree", "Master's degree", "Doctor's degree")))

# View the plot again
p

# 2. X-axis labels need to include the whole academic year
# Adjust the scale for the x-axis
# It might be nice to also use commas in the y-axis labels
# Adjust the scale for the y-axis
p_1 <- p +
  scale_x_continuous(breaks = seq(2017, 2020, 1),
                     labels = c("2016-17", "2017-18", "2018-19", "2019-20")) +
  scale_y_continuous(labels = comma)

p_1

# 3. Axis titles are not very helpful
# 4. There's not plot title or data source noted
# Add on the labs() function
p_2 <- p_1 +
  labs(title = "Number of Degrees Awarded by the University of Arizona",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data Source: IPEDS Completions Survey")

p_2

# 5. Use colors other than the R defaults for the lines
# This is another application of a scale function
p_3 <- p_2 +
  scale_color_manual(values = c("#0C234B", "#AB0520", "#378DBD")) 

p_3

# We can also use other pre-built color schemes 
p_2 +
  scale_color_brewer(palette = "Dark2") 

# 6. Remove the grey background from the plot
# And move the legend to the top
p_4 <- p_3 +
  theme_minimal() +
  theme(legend.position = "top")

p_4

# Final touch-ups
# Add a point for each year
p_5 <- p_4 + 
  geom_point()

p_5

# To remove the line in the legend, we can go back to the beginning and change geom_line() to geom_line(show.legend = FALSE)

# Save the plot
ggsave(filename = here("plots", "line_plot.png"),
       plot = p_5,
       device = "png",
       width = 8,
       height = 5,
       type = "cairo")

```

## Fixing Potential Errors

```{r data-viz-08-erro, eval = FALSE}

# Why doesn't the code in this chunk create what we expect?

# Get the data ready for the plot
ipeds_adm_plot <- ipeds_adm %>%
  filter(year == 2019)

## 1 -  Color ##

# Not quite right
ggplot(data = ipeds_adm_plot,
       mapping = aes(x = sat_score,
                     y = act_score,
                     color = "blue")) +
  geom_point()


# Correct
ggplot(data = ipeds_adm_plot,
       mapping = aes(x = sat_score,
                     y = act_score)) +
  geom_point(color = "blue")

## 2 - Plus sign ##

# Not quite right
ggplot(data = ipeds_adm_plot,
       mapping = aes(x = sat_score,
                     y = act_score)) 
 + geom_point() # <- + at the beginning


ggplot(data = ipeds_adm_plot,
       mapping = aes(x = sat_score,
                     y = act_score)) # <- missing +
  geom_point()
  
ggplot(data = ipeds_adm_plot,
       mapping = aes(x = sat_score,
                     y = act_score)) + + # <- double +
  geom_point()

# Correct
ggplot(data = ipeds_adm_plot,
       mapping = aes(x = sat_score,
                     y = act_score)) + # Need to have a single + at the end of the line
  geom_point()


## 3 - Missing object ##

# Not quite right
ggplot(data = ipeds_adm_plot,
       mapping = aes(x = sat,
                     y = act_score)) +
  geom_point()

# Correct
ggplot(data = ipeds_adm_plot,
       mapping = aes(x = sat_score,
                     y = act_score)) +
  geom_point()

## 4 - Parentheses ##

# Not quite right
ggplot(data = ipeds_adm_plot,
       mapping = aes(x = sat_score,
                     y = act_score) + # <- missing parentheses
  geom_point()
  
ggplot(data = ipeds_adm_plot,
       mapping = aes(x = sat_score,
                     y = act_score))) + # <- extra parentheses 
  geom_point()

# Correct
ggplot(data = ipeds_adm_plot,
       mapping = aes(x = sat_score,
                     y = act_score)) +
  geom_point()

```


## Data Visualization - Your Turn

Using the code we've already written, try making a new version of a plot. Use a different geometric object (i.e. replace `geom_line()` with `geom_bar()`) or change the aesthetics mappings (i.e. add a color or fill argument to the `aes()` function or switch out the variables used in the `aes()` function to visualize different values). 

Make a visualization using a different IPEDS dataset. 

Learn more at the package website: https://ggplot2.tidyverse.org/ 


```{r visualize-data-your-turn}




```


