---
title: "Data Analysis and Visualization in R"
author: "Jenn Schilling"
date: "June 5, 2022"
output:
 html_document:
  theme: "flatly" 
  toc: TRUE
  toc_float: TRUE
---


This file contains contains code for pre-conference AIR Forum Workshop: Data Analysis Visualization in R, presented for the Association for Institutional Research by Jenn Schilling on June 5, 2022. 

**Materials developed by Jenn Schilling.**  


```{r setup, include=FALSE}

# Set default chunk options
knitr::opts_chunk$set(echo = TRUE, message = FALSE) 

# Set the default size of figures
knitr::opts_chunk$set(fig.width = 8, fig.height = 5)  

# Load libraries
library(here)
library(explore)
library(tidyverse)
library(scales)

```


# Read the Data

```{r read-data, include=FALSE}

# Data from IPEDS
ipeds_char <- read_csv(here("data", "IPEDS_inst_char.csv"))
ipeds_adm <- read_csv(here("data", "IPEDS_inst_adm.csv"))
ipeds_enrl <- read_csv(here("data", "IPEDS_inst_enrl.csv"))
ipeds_compl <- read_csv(here("data", "IPEDS_inst_compl.csv"))

```



# Data Exploration

```{r explore-data}

# Interactive exploration of the data
explore(ipeds_adm)

# Create a report for exploratory data analysis
report(ipeds_adm,
       output_file = "ipeds_adm_report.html",
       output_dir = here("output"))

# Create a report for exploratory data analysis with a target variable
report(ipeds_adm,
       output_file = "ipeds_adm_report_target.html",
       output_dir = here("output"),
       target = n_enrl)

```


## Your Turn

Use the `explore()` or `report()` to explore one of the other IPEDS datasets.

What do you notice? 

Are there any outliers or unusual patterns in any variable?

Are there any missing values? 

Does the shape of the data reveal anything interesting? 

```{r explore-data-your-turn}




```



# Data Analysis

```{r analyze-data}

# Summarize Data


# Manipulate Data (pivot, join, mutate, filter)


# Tidy Data


```


## Your Turn

```{r analyze-data-your-turn}




```



# Data Visualization


## Your Turn


# From Previous AIR Webinar

# Scatter plot

Make a scatter plot 

```{r scatter-plot}

inst_adm_plot <- left_join(inst_list, inst_adm, by = "unitid") %>%
  filter(year == 2019)

ggplot(data = inst_adm_plot,
       mapping = aes(x = sat_score,
                     y = act_score)) +
  geom_point()

```

# Line Graph

Make a line graph

```{r line-graph}

inst_adm_plot <- left_join(inst_list, inst_adm, by = "unitid")

ggplot(data = inst_adm_plot,
       mapping = aes(x = year,
                     y = sat_score)) +
  geom_line()

# Something went wrong

# Need to add group mapping
ggplot(data = inst_adm_plot,
       mapping = aes(x = year,
                     y = sat_score,
                     group = instnm)) +
  geom_line()

# What about number who submitted by year?
ggplot(data = inst_adm_plot,
       mapping = aes(x = year,
                     y = n_sat,
                     group = instnm)) +
  geom_line()

# Who's that outlier in 2019?
ggplot(data = inst_adm_plot,
       mapping = aes(x = year,
                     y = n_sat,
                     group = instnm)) +
  geom_line() +
  geom_text(mapping = aes(label = instnm)) # Add Text Labels


# We can see it's Penn State, but the labels are kind of a mess
ggplot(data = inst_adm_plot,
       mapping = aes(x = year,
                     y = n_sat,
                     group = instnm)) +
  geom_line() +
  # Filter to only the last year for the text labels
  geom_text(data = inst_adm_plot %>% filter(year == 2019),
            mapping = aes(x = year,
                          y = n_sat,
                          label = instnm))

# How can we make the labels more readable?
ggplot(data = inst_adm_plot,
       mapping = aes(x = year,
                     y = n_sat,
                     group = instnm)) +
  geom_line() +
  # Change the x mapping for the text labels and add some format adjustments 
  geom_text(data = inst_adm_plot %>% filter(year == 2019),
            mapping = aes(x = Inf,
                          y = n_sat,
                          label = instnm),
            hjust = -.01) +
  coord_cartesian(clip = "off") +
  theme(plot.margin = margin(0, 250, 0, 0))


```

# Bar Graph

Make a bar graph

```{r bar-plot}

# How many institutions of each type?
ggplot(data = inst_list,
       mapping = aes(x = locale)) +
  geom_bar()

# Can we sort them?
# Yes, by using the fct_infreq() function from the forcats package in tidyverse
# fct_infreq() reorders the levels by their frequencies
ggplot(data = inst_list,
       mapping = aes(x = fct_infreq(locale))) +
  geom_bar()

# What if we want to make a bar plot with something other than counts?

inst_enrlmnt_plot <- left_join(inst_list, inst_enrlmnt, by = "unitid")

# What does enrollment look like by urban area?

# This will give total enrollment in each urban area
ggplot(data = inst_enrlmnt_plot %>% 
         filter(lstudy == "All students", gender == "Total", year == 2019),
       mapping = aes(x = locale,
                     y = n_enrl)) +
  geom_col()

# See - the enrollments for each institution are stacked on top of one another
ggplot(data = inst_enrlmnt_plot %>% 
         filter(lstudy == "All students", gender == "Total", year == 2019),
       mapping = aes(x = locale,
                     y = n_enrl,
                     fill = instnm)) +
  geom_col()

# What if we want the median enrollment in each area? 
inst_enrlmnt_plot_sub <- inst_enrlmnt_plot %>%
  filter(lstudy == "All students", gender == "Total", year == 2019) %>%
  group_by(locale) %>%
  summarize(med_enrl = median(n_enrl),
            .groups = "drop")

ggplot(data = inst_enrlmnt_plot_sub,
       mapping = aes(x = locale,
                     y = med_enrl)) +
  geom_col()

# Sort by median enrollment
# This time, use the reorder() function to arrange the locale column by the med_enrl column
ggplot(data = inst_enrlmnt_plot_sub,
       mapping = aes(x = reorder(locale, -med_enrl),
                     y = med_enrl)) +
  geom_col()

# What if we want to see the enrollment by institution and urban area?
ggplot(data = inst_enrlmnt_plot %>% 
         filter(lstudy == "All students", 
                gender == "Total",
                year == 2019),
       mapping = aes(x = n_enrl,
                     y = reorder(instnm, 
                                 n_enrl))) +
  geom_col() +
  facet_wrap(~ locale,
             scales = "free_y",
             ncol = 1)

```

# Add color

Add color to bar graph

```{r fill-color}

inst_enrlmnt_plot <- left_join(inst_list, inst_enrlmnt, by = "unitid")

# How has median enrollment changed by urban area over time?
inst_enrlmnt_plot_sub <- inst_enrlmnt_plot %>%
  filter(lstudy == "All students", gender == "Total") %>%
  group_by(year, locale) %>%
  summarize(med_enrl = median(n_enrl),
            .groups = "drop")

ggplot(data = inst_enrlmnt_plot_sub,
       mapping = aes(x = year,
                     y = med_enrl,
                     fill = locale)) +
  geom_col() 


# Remove the stacking
ggplot(data = inst_enrlmnt_plot_sub,
       mapping = aes(x = year,
                     y = med_enrl,
                     fill = locale)) +
  geom_col(position = "dodge") 

# Make a different plot for each urban area
ggplot(data = inst_enrlmnt_plot_sub,
       mapping = aes(x = year,
                     y = med_enrl,
                     fill = locale)) +
  geom_col(position = "dodge") +
  facet_wrap(~ locale) +
  guides(fill = FALSE) # remove the legend

# Maybe it would be better to show this with lines?
ggplot(data = inst_enrlmnt_plot_sub,
       mapping = aes(x = year,
                     y = med_enrl,
                     fill = locale)) +
  geom_line() +
  facet_wrap(~ locale) +
  guides(fill = FALSE) 

# What happened to the color?
# Bars, histograms, and boxplots use fill; lines and points use color
ggplot(data = inst_enrlmnt_plot_sub,
       mapping = aes(x = year,
                     y = med_enrl,
                     color = locale)) +
  geom_line() +
  facet_wrap(~ locale) +
  guides(color = FALSE) 

```


# Labels and Formatting

Add labels and formatting


```{r formatting}

inst_compl_plot <- left_join(inst_list, inst_compl, by = "unitid") 

ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", 
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = awlevelc)) +
  geom_line()

# Reorder degree levels
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", 
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line()

# Add titles and labels
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total",
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line() +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "")

# Move legend to the top 
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", 
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line() +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "") +
  theme(legend.position = "top")

# Add note about data source
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", 
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line() +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme(legend.position = "top")


# Add Theme
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", 
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line() +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme_minimal() +
  theme(legend.position = "top")


# Add commas to y-axis labels
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", 
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme_minimal() +
  theme(legend.position = "top")

# Remove decimals from x-axis labels and make them academic year ranges
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", 
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = c(2017, 2018, 2019)) +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme_minimal() +
  theme(legend.position = "top")

ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", 
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = c(2017, 2018, 2019),
                     labels = c("2016-17", 
                                "2017-18", 
                                "2018-19")) +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme_minimal() +
  theme(legend.position = "top")


# Add points on top of the lines and remove line legend
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total",
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line(show.legend = FALSE) +
  geom_point() +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = c(2017, 2018, 2019),
                     labels = c("2016-17", 
                                "2017-18", 
                                "2018-19")) +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme_minimal() +
  theme(legend.position = "top")

# Change color using ColorBrewer
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total",
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line(show.legend = FALSE) +
  geom_point() +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = c(2017, 2018, 2019),
                     labels = c("2016-17", 
                                "2017-18", 
                                "2018-19")) +
  scale_color_brewer(palette = "Dark2") +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme_minimal() +
  theme(legend.position = "top")

# Change color using manual colors
ggplot(data = inst_compl_plot %>% 
         filter(gender == "Total", 
                unitid == 104179),
       mapping = aes(x = year,
                     y = n_deg,
                     color = factor(awlevelc,
                      levels = c("Bachelor's degree",
                                 "Master's degree",
                                 "Doctor's degree")))) +
  geom_line(show.legend = FALSE) +
  geom_point() +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = c(2017, 2018, 2019),
                     labels = c("2016-17", 
                                "2017-18", 
                                "2018-19")) +
  scale_color_manual(values = c("#0C234B", 
                                "#AB0520", 
                                "#378DBD")) +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Survey") +
  theme_minimal() +
  theme(legend.position = "top")


```


# Saving

Save a plot

```{r save}

inst_compl_plot <- left_join(inst_list, inst_compl, by = "unitid") 

line_plot <- ggplot(data = inst_compl_plot %>% 
                      filter(gender == "Total", 
                             unitid == 104179),
                    mapping = aes(x = year,
                                  y = n_deg,
                                  color = factor(awlevelc,
                                                 levels = c("Bachelor's degree",
                                                            "Master's degree",
                                                            "Doctor's degree")))) +
  geom_line(show.legend = FALSE) +
  geom_point() +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = c(2017, 2018, 2019),
                     labels = c("2016-17", "2017-18", "2018-19")) +
  scale_color_manual(values = c("#0C234B", "#AB0520", "#378DBD")) +
  labs(title = "Number of Degrees Awarded by the University of Arizona by Academic Year",
       x = "Academic Year",
       y = "",
       color = "",
       caption = "Data from IPEDS Completions Surveys") +
  theme_minimal() +
  theme(legend.position = "top")

ggsave(filename = here("plots", "line_plot.png"),
       plot = line_plot,
       device = "png",
       width = 8,
       height = 5,
       type = "cairo")


```

